A new metric for automatic discovery of periodic patterns in time series

### 2021-02-28 paper/main.Rmd
notations and characterising distribution section re-written

@@

Each of the data subsets in the data structure have multiple observations and may vary widely across different subsets due to the structure of the calendar, missing observations or uneven locations of events in the time domain. The set of observations corresponding to each combination is assumed to be a sample from an unknown probability density function.
While the whole population of observations has certain characteristics, we can typically never measure all of them. Often shape, central tendency, and variability are the common characteristics used to describe the distribution. Another way to describe the probability distribution is through quantiles. (Define quantiles here) <!--There are two broad approaches to quantile estimation, viz, parameteric and non-parameteric. The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data.--> Sample quantiles could be thought to estimate the population quantiles. But there are a large number of different definitions used for sample quantiles. The median-unbiased estimator is recommended (Rob's paper) because of its desirable properties of a quantile estimator and can be defined independently of the underlying distribution.

>>>>

## Notations

Consider two cyclic granularities $A$ and $B$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets. Let $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ be a continuous variable observed across $T$ time points. Let the four elementary designs as described in Figure \ref{fig:null4by2} be $D_{null}$ where there is no difference in distribution of $v$ for $A$ or $B$, $D_{var_f}$ denotes the set of designs where there is difference in distribution of $v$ for $B$ and not for $A$. Similarly, $D_{var_x}$ denotes the set of designs where difference is observed only across A. Finally, $D_{var_{all}}$ denotes those designs for which difference is observed across both $A$ and $B$.

@@

Multiple observations of $v$ correspond to the subset  $v_{jk} = \{s: A(s) = j, B(s) = k\}$. The number of observations and the structure might vary widely across subsets due to the structure of the calendar, missing observations or uneven locations of events in the time domain. Each $v_{jk} : j\in \{1, 2, \dots, J\}, k\in \{1, 2, \dots, K\}\}$ is assumed to be drawn from a continuous probability distribution and have certain characteristics. Often shape, central tendency, and variability are the common characteristics used to describe the distribution. Mean, median or mode are generally used to describe the center of the distribution, while range, standard deviation, quantiles, standard errors and confidence intervals are often used to describe variability in the distribution. Quantiles are chosen as a way to characterize distributions in this paper.

The quantile of a distribution with probability $p$ is defined as $Q(p)=F^{-1}(p) = inf\{x: F(x) >p\}$, $0<p< 1$ where $F(x)$ is the distribution function. There are two broad approaches to quantile estimation, viz, parameteric and non-parameteric. The benefit of using a non-parametric estimator is that there are less rigid assumptions made about the nature of the underlying distribution of the data. Sample quantiles could be used for estimation in a non-parametric setup. @Hyndman1996-ty describes the difficulty in defining sample quantiles and recommends the use of median-unbiased estimator because of "desirable properties of a quantile estimator and can be defined independently of the underlying distribution.". The `stats::quantile()` function in @R-language could be used for practical implementation where type = 8 refers to the algorithm corresponding to the median-unbiased estimator.
